---
title: "Business Insights Report"
subtitle: "Data Visualization and Communication"
author:
  - "Mia--667174866"
  - "Pavlina--667753859"
date: "2025-10-15"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: false
---

# Executive Summary

This report analyzes the e-commerce company’s quarterly performance data. Insights include the highest performing sales channels, average customer trends per quarter, and revenue efficiency per customer. Visualizations reveal which channels drive the most revenue and convert customers efficiently. Based on these analyses, strategic recommendations prioritize investment in high-performing channels to maximize revenue and profitability. Specifically, the online channel generates the highest total revenue ($5,571,130), while Retail achieves the highest revenue per customer ($314.22), which highlights which channels are clearly more profitable overall and per customer. 

## Question 1: Data Preparation and Initial Analysis

### a) Load and Explore the Data


```{r}
# Load ggplot2
library(ggplot2)

# Load the business dataset
business_data <- read.csv("business_performance.csv")

# Inspect structure and summary
str(business_data)
summary(business_data)
```

Answer (a) : From the output, both quarter and channel are stored as character variables. This means that they represent text data rather than numeric values. The revenue variable ranges from a minimum of 150,231 to a maximum of 412,125, indicating the lowest and highest revenue values recorded in the dataset. 

### b) calculate total revenue by channel

```{r}
aggregate(revenue ~ channel, 
          data = business_data,
          FUN = sum
          )

aggregate(customers ~ quarter,
           data = business_data,
           FUN = mean
           )

```

Answer (b) : Based on aggregated results, the Online channel generates the highest total revenue overall, with the total of 5,571,130, which is followed by Mobile and Retail. When seeking out the average number of customers by quarter, Quarter 3 (Q3) has the most customers on average at approximately 1,172.9, indicating that this is in fact its busiest quarter. 


### c) 

```{r}
business_data$revenue_per_customer <- business_data$revenue / business_data$customers

aggregate(revenue_per_customer ~ channel , data = business_data , mean,  na.rm = TRUE )
```

Answer (c) : After calculating the revenue per customer and then averaging that number by channel, the Retail channel has the highest average revenue per customer at approx. $314.22. What this suggests is that although Retail may generate less total revenue than Online, each individual customer in the Retail channel tends to spend more on an average, which indicates higher-value transactions per customer. 



## 2 Creating Professional Charts


### a)

```{r}
# Calculate totals first
channel_totals <- aggregate(revenue ~ channel, data = business_data, sum)
# Then create the chart
ggplot(channel_totals, aes(x = channel, y = revenue)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Revenue by Channel", 
       x = "Sales Channel", y = "Revenue ($)")
```

Answer (a) : Based on the bar chart showing total revenue by channel, the Online channel generates the highest overall revenue. This makes it apparent that the company should in fact consider allocating more investment into the Online channel, because of the fact that it appears to be the strongest performer, clearly offering the greatest potential for continued profitability and overall growth. 

### b) 

```{r}
ggplot(business_data, aes(x = customers, y = revenue, color = channel)) +
  geom_point(size = 3) +
  labs(title = "Customer Count vs Revenue by Channel",
       x = "Number of Customers", y = "Revenue ($)")
```

Answer (b) : Based on the visualization output, the Retail channel appears to be the most eficient at converting customers into revenue. Although it has FEWEST customers, it still generates relatively high amounts of revenue compared to the other channels. What this suggests is that Retail produces more revenue per customer on average, indicating overall stronger efficiency in driving higher sales from a smaller customer base. 


## 3 Action Plan

### a) 

```{r}
# Load knitr for clean tables
library(knitr)

# 1) Total revenue by channel
totals <- aggregate(revenue ~ channel, data = business_data, sum)

# 2) Average customers by channel
avg_customers <- aggregate(customers ~ channel, data = business_data, FUN = mean, na.rm = TRUE)

# 3) Average revenue per customer
business_data$revenue_per_customer <- business_data$revenue / business_data$customers
avg_rpc <- aggregate(revenue_per_customer ~ channel, data = business_data, FUN = mean, na.rm = TRUE)

# 4) Average conversion rate by channel
avg_conversion <- aggregate(conversion_rate ~ channel, data = business_data, FUN = mean, na.rm = TRUE)

# 5) Merge all metrics into a single table
summary_table <- merge(totals, avg_customers, by = "channel")
summary_table <- merge(summary_table, avg_rpc, by = "channel")
summary_table <- merge(summary_table, avg_conversion, by = "channel")

# 6) Rename columns for presentation
colnames(summary_table) <- c("Channel", "Total Revenue", "Average Customers", 
                             "Average Revenue per Customer", "Average Conversion Rate")

# 7) Display a professional table in R Markdown
kable(summary_table, digits = 2, caption = "Summary of Key Metrics by Channel")
```



Answer 3 (b) : 

Invest additional resources in the Online channel, which is what generates the highest total revenue of $5,571,130 and expands the largest customer base (1,559.56 customers on average). While Retail achieves the highest revenue per customer ($314.22), its smaller audience limits impact overall. Prioritizing Online marketing, customer experience, technology enhancements, and retention initiatives will maximize overall revenue growth and profitability in the long-term while maintaining overall efficiency in customer acquisition across all of the channels. 
